Classical probability measures as pullback of quantum probability measures?

### 2022-04-09 pullback.tex
pullback.lyx change the main compiler from XeTeX to pdfLaTeX...

%% LyX 2.3.2-2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt]{iuphd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{rotfloat}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{url}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[all]{xy}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\providecommand{\LyX}{L\kern-.1667em\lower.25em\hbox{Y}\kern-.125emX\@}
\let\@@LyX\LyX
\def\LyX{\@ensure@LTR{\@@LyX}}
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newlength{\lyxlabelwidth}      % auxiliary length 
\providecommand*{\code}[1]{\texttt{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{csquotes}
\usepackage{comment}
\usepackage{framed}
\usepackage{mathrsfs}
\usepackage{setspace}
\usepackage{dsfont}
\usepackage[AutoFallBack=true]{xeCJK}
\setCJKmainfont[FallBack=Batang]{SimSun}
% \usepackage{xeCJK}
% \setCJKmainfont{SimSun}

@@

%\usepackage[basic]{complexity}
\newcommand{\lang}[1]{{\ensuremath{\textsf{#1}}}}
\newcommand{\newlang}[2]{\newcommand{#1}{\lang{#2}}}
\newlang{\usat}{UNIQUE-SAT}
\newcommand{\ComplexityFont}[1]{%
{\ensuremath{\textsf{#1}}}%extra {} makes everyone happy.
}
\newcommand{\NP}{\ComplexityFont{NP}}

\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{prop}{Proposition}[chapter]
\newtheorem{thm}{Theorem}[chapter]
\newtheorem{lemma}{Lemma}[chapter]
\newtheorem{case}{Case}
\newtheorem{corollary}{Corollary}[chapter]

\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{question}{Question}[chapter]
\newtheorem*{method}{Method}
\newtheorem{example}{Example}[chapter]
\newtheorem{postulate}{Postulate}[chapter]

@@

\usepackage{unicode-math}
\RequirePackage{ifthen}
\@ifpackageloaded{unicode-math}{}{
 \newcommand{\symbf}{\mathbf}
 \newcommand{\symrm}{\mathrm}
 \newcommand{\symcal}{\mathcal}
 \newcommand{\symbb}{\mathbb}
}
\newcommand{\muB}{\ensuremath{\mu^{\symrm{B}}}}
\newcommand{\muC}{\ensuremath{\mu^{\symrm{C}}}}
\newcommand{\muF}{\ensuremath{\mu^{\symrm{F}}}}
\newcommand{\symrmL}{\ensuremath{\symrm{L}}}
\newcommand{\symrmR}{\ensuremath{\symrm{R}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\symrmL{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\symrmR{#1}}}}
\newcommand{\barmuD}{\ensuremath{\bar{\mu}^{\symrm{D}}}}
\newcommand{\events}{\ensuremath{\symcal{E}}}
\newcommand{\eventsC}{\ensuremath{\events_{\symrm{C}}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\symbb{F}_{{#1}}}
\newcommand{\Fp}{\Fpx{p}}
\newcommand{\Fppx}[1]{\Fpx{{#1}^2}}
\newcommand{\Fpp}{\Fppx{p}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\symbf{#1}}}
\newcommand{\mathReal}{\symbb{R}}
\newcommand{\mathComplex}{\symbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\symbf{S}^{#1}}
\newcommand{\R}{\mathReal}
\newcommand{\rme}{\symrm{e}}
\newcommand{\rmi}{\symrm{i}}
\newcommand{\rmd}{\symrm{d}}
\newcommand{\CP}[1]{\mathComplex\symbf{P}^{#1}} % Complex projective space
\newcommand{\DCP}[1]{\symbf{D}\mathComplex\symbf{P}^{#1}} % Discrete complex projective
\def\fh{\mathfrak{h}}
\newcommand{\uf}{U_{\!f}}
\newcommand{\scalarPlus}{+}
\newcommand{\boolt}{\textsf{Bool}} 
\newcommand{\bfalse}{\texttt{\textbf{false}}}
\newcommand{\btrue}{\texttt{\textbf{true}}}
\newcommand{\dotprod}{dot product}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\newcommand{\Psibar}{\overline{\Psi}}
\newcommand{\cardp}[2]{#1 \, \slash\!\! \slash \, #2}
\def\round{\mathop{\rm round}\nolimits}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\Hilb}{\symcal{H}}
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\def\C{{\symbb{C}}}
\newcommand{\ultramodular}{\symcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\frameL}[1][]{\ensuremath{f^{L{#1}}}}
\newcommand{\frameR}[1][]{\ensuremath{f^{R{#1}}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\nb}{\nolinebreak[3] }
\NewDocumentCommand{\jthSubsystem}{m O{} m}{{#1}_{#2}^{#3}}

@@

\department{Department of Mathematics}
\department{Department of Computer Science}

\usepackage{datetime2}
\usepackage{datetime2-calc}
\monthGranted{June}%\DTMmonthname{\month}}
\yearGranted{\the\year}

@@

\renewcommand{\tabularxcolumn}[1]{m{#1}}
\newcolumntype{Y}{>{\centering\doublespacing\arraybackslash}X}
\newcolumntype{M}[1]{>{\centering\doublespacing\arraybackslash}m{\widthof{#1}}}

@@

\@ifundefined{showcaptionsetup}{}{%
 \PassOptionsToPackage{caption=false}{subfig}}
\usepackage{subfig}
\makeatother

\usepackage{polyglossia}
\setdefaultlanguage[variant=american]{english}
\setotherlanguage{arabic}
\setotherlanguage{bahasai}
\setotherlanguage{basque}
\setotherlanguage{french}
\setotherlanguage{german}
\setotherlanguage{italian}
\setotherlanguage{portuges}
\setotherlanguage{turkish}
\setotherlanguage{vietnamese}
\usepackage[style=numeric-comp,maxbibnames=10,doi=false,isbn=false,url=false]{biblatex}
\addbibresource{prop.bib}
\begin{document}
\title{Discrete Quantum Theories and Computing}
\author{Yu-Tsung Tai}

\maketitle
%data for Acceptance Page
\committeeMember{Amr A. Sabry, PhD} \committeeMember{Dylan Paul
Thurston, PhD} \committeeMember{Gerardo Ortiz, PhD} \committeeMember{Andrew
J. Hanson, PhD} \committeeMember{Shouhong Wang, PhD} \defensedate{November
29, 2018} \acceptancepage

%data for Copyright Page
%\copyrightpage

\begin{dedication}To my parents, Cheng-Tien Tai (戴振沺) and Feng-Ming
Chang (張鳳鳴).\end{dedication}

\chapter*{Survey of the acknowledgments in theses}

I found the sentences in my Acknowledgments almost always contains
“thank” and looks boring. To improve my Acknowledgments, I briefly
read the Acknowledgments in other theses \cite{Chih2013,Sun2015,Lee2013,Chen2016,sabry1995formal}.
\begin{acknowledgments*}

@@

\noindent \centering{}{\small{}\caption{\label{tab:committee}The professors served in all my committees.}
}%% ERT block 1 (before LyX tabular)

@@

{\small{}}%
\begin{tabular}{M{\small{}MATH-M781}M{\small{}Co-author}M{\small{}(\textarabic{عمرو~صبري})}M{\small{}Co-author}M{\small{}Thurston}M{\small{}Shouhong}M{\small{}Lawrence}Y}

@@

{\small{}Committee or course} & {\small{}Gerardo Ortiz} & {\small{}Amr\nolinebreak[3] A. Sabry\linebreak[3] (}\textarabic{{\small{}عمرو
صبري}}{\small{})} & {\small{}Andrew\nolinebreak[3] J. Hanson} & {\small{}Dylan\nolinebreak[3] Paul Thurston} & {\small{}Shouhong Wang\linebreak[3] (}{\small{}汪守宏}{\small{})} & {\small{}Lawrence\nolinebreak[3] S.\linebreak[2] Moss} & {\small{}Relation with the dissertation}\tabularnewline

@@

 & {\small{}Co-author} & {\small{}Co-author} & {\small{}Co-author} &  &  &  & {\small{}Most of Chapters~\ref{chap:Conventional-Quantum-Theory},
\ref{chap:Unrestricted-Finite-Fields}, and~\ref{chap:DQT=000026DQC}
is based on~\cite{geometry2013} and~\cite{DQT2014}; most of Chapter~\ref{chap:QIVPM}
is based on~\cite{THOS2018}.}\tabularnewline

@@

{\small{}Research committee} & {\small{}Co-chair} & {\small{}Co-chair} & {\small{}Member} & {\small{}Co-chair} & {\small{}Member} &  & {\small{}Some questions in the dissertation proposal are answered
in Chapter~\ref{chap:QIVPM}, and some of the remaining ones are
explained in Chapter~\ref{chap:Further-Discussion}.}\tabularnewline

@@

 &  & {\small{}Supervisor} &  &  &  &  & {\small{}The application for \emph{{\small{}Rethinking Foundations of Physics}}
2017 workshop inspires Chapter~\ref{chap:Introduction}.}\tabularnewline

@@

{\small{}Advisory committee} & {\small{}Member} & {\small{}Member} & {\small{}Member} & {\small{}Member} &  &  & {\small{}The survey of real computation in the Computer Science qualifying
exam (chaired by Prof.\ Sabry) inspires Chapter~\ref{chap:Introduction}.}\tabularnewline

@@

{\small{}PHYS-P700} & {\small{}Instructor} &  &  &  &  &  & {\small{}The textbook~\cite{Mermin2007} of \emph{{\small{}Quantum Computation and Information}}
is cited in Chapters~\ref{chap:Introduction}, \ref{chap:Conventional-Quantum-Theory},
\ref{chap:DQT=000026DQC}, and~\ref{chap:QIVPM}.}\tabularnewline

@@

{\small{}CSCI-B629} &  & {\small{}Instructor} &  &  &  &  & {\small{}The textbook~\cite{hottbook2013} of \emph{{\small{}Homotopy Type Theory}}
is cited in Sec.~\ref{discretequantumtheoryIIb}.}\tabularnewline

@@

{\small{}Tier 3 committee} & {\small{}Co-chair} & {\small{}Co-chair} &  & {\small{}Member} &  & {\small{}Co-chair} & {\small{}Two of the assigned papers \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644}
in the Tier 3 exam~\cite{Mathematics2018} are cited in Chapters~\ref{chap:Introduction}
and~\ref{chap:Unrestricted-Finite-Fields}.}\tabularnewline

@@

{\small{}MATH-M781} &  &  &  &  &  & {\small{}Instructor} & {\small{}The presented paper~\cite{Abramsky2012} of \emph{{\small{}Coalgebra}}
is cited in Sec.~\ref{subsec:Classical-and-Quantum}.}\tabularnewline

@@

I want to thank the professors who served in all my committees. These
professors advised me as committee members and taught me while we
wrote papers together and when I took their courses. Most of my dissertation
is either based on these papers or inspired by the studies supervised
by them as listed in Table~\ref{tab:committee}. Especially, Prof.\ Andrew
J. Hanson, Prof.\ Gerardo Ortiz, and Prof.\ Amr A. Sabry\ (\textarabic{عمرو صبري})
meet me weekly for the past few years, read and tried to understand
what I typed which sometimes even myself cannot understand, taught
me how to organize them in ``English'' so that general audience
might be interested and have some chance to understand, and encouraged
me to participate seminars, workshops, and conferences. Prof.\ Lawrence
Moss invited me to present our results in the interdisciplinary logic
seminar and theory seminar in Computer Science and wrote an assessment
letter for me. Although not directly related to this dissertation,
the \textgerman[variant=german,spelling=new,babelshorthands=true]{Weil}
conjecture~\cite{hartshorne1977algebraic} on Grassmannians studied
with Prof.\ Dylan Paul Thurston and the quantum interpretation discussed
with Prof.\ Shouhong Wang (汪守宏) also broadened my understanding on
discrete and conventional quantum theories. I would like to express
my appreciation to all of these.

To have the current content of dissertation, I would also like to
thank John Gardiner for some inspiring discussion about quantum probability
measures over finite fields~\cite{Gardiner2014}. His research supervised
by Prof.\ Gerardo Ortiz inspired Sec.~\ref{subsec:Maximal-entanglement}
and~\ref{sec:Toward-IVPM}. Besides of my committee members, what
I learned from the teachers in other courses also helped me consolidate
this dissertation as listed in Table~\ref{tab:teacher}. Learning
other materials also made me more mature and easier to communicate
with others. I need to thank those who taught me, especially my English
teachers, Traci Nagle, Elizabeth ``Betsy'' Merceron, and Kexin ``Casey''
Chen (谌可心).

\begin{table}[!t]
\begin{doublespace}
\noindent \centering{}\caption{\label{tab:teacher}Other courses I took which is relevant to this
dissertation, where IUB stands for Indiana University Bloomington
and NTU stands for National Taiwan University.}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{M{Chou~(周謀鴻)}cM{MATH-M522}Y}
\toprule 
Instructor & School & Course & Relation with the dissertation\tabularnewline
\midrule
\midrule 
Tom Lewis & IUB & SLST-T501 & Sec.~\ref{subsec:Discrete-Deutsch-algorithm} is based on the term
paper of \emph{Academic Writing}.\tabularnewline
\midrule 
Michael\nolinebreak[3] A. Mandell & IUB & MATH-M522 & The textbook~\cite{Hatcher2001} of \emph{Topology II} is cited
in Chapters~\ref{chap:Introduction} and~\ref{chap:Conventional-Quantum-Theory}.\tabularnewline
\midrule 
Steven Myers & IUB & CSCI-B502 & The textbook~\cite{AroraBarak2009} of \emph{Computational Complexity}
is cited in Sec.~\ref{modalquantumcomputing}.\tabularnewline
\midrule 
Valery Lunts & IUB & MATH-M501 & The textbook~\cite{DummitFoote2004} of \emph{Survey of Algebra}
is cited in Secs.~\ref{sec:background}, \ref{discretequantumtheoryI},
and~\ref{discretequantumtheoryIIb}.\tabularnewline
\midrule 
Huah Chu\linebreak[3] (朱樺\textenglish[variant=american]{)} & NTU & 221~U3830\linebreak[3] 221~U3840 & The textbook~\cite{Artin1991} of \emph{Algebra (I)} and \emph{(II)}
is cited in Secs.~\ref{subsec:Two-dimensional-Hilbert-Space}, \ref{subsec:Cyclic-Properties-of},
\ref{subsec:Complexified-Finite-Fields}, and~\ref{discretequantumtheoryIIb}.\tabularnewline
\midrule 
Jin-Tzu Chen\linebreak[3] (陳金次) & NTU & 201~31300 & The textbook~\cite{GAMELIN2003} of \emph{Functions of a Complex Variable}
is cited in Sec.~\ref{subsec:Vector-Spaces}.\tabularnewline
\midrule
Mo-Hong\linebreak[3] Chou (周謀鴻) & NTU & 221~U4290 & The textbook~\cite{GolubVanLoan1996} of \emph{Introduction to Computational Linear Algebra}
is cited in Sec.~\ref{subsec:Finite-Precision-Extension-of}.\tabularnewline
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{doublespace}
\end{table}
Many people have done everything to make my double-major smooth. I
need to thank all of their supports, financially or providing some
experience, as listed in my Curriculum Vitae at the end of the dissertation.
Most importantly, Prof.\ Amr A. Sabry\ (\textarabic{عمرو صبري})
helped me deal with the rules, supervised my reading courses for writing
our papers, and financially support me as a research assistant and
associate. In the department of Mathematics, I need to thank the directors
of graduate study in Mathematics, especially Prof.\ Christopher Judge,
Prof.\ Matthias Weber, and Prof.\ Michael A.\ Mandell; and the
graduate services assistant, Kate Forrest. In the department of Computer
Science, I want to thank the directors of \PhD study in Computer
Science, especially Prof.\ Yuqing Melanie Wu (吴愈青) and Prof.\ \textturkish{Funda
Ergun}; the graduate student office; the director of AI/UI assignments,
Charles Pope; and the infrastructure and technology group (ITG). Particularly,
I don't know how much computational power I consumed to count the
states in Sec.~\ref{subsec:Maximal-entanglement}. This computational
power not only came from the Linux systems in Computer Science but
also came from the Big Red II administrated by University Information
Technology Services (UITS). Indiana University Libraries was also
extremely helpful during my dissertation research. Except for a stack
of books I checked out directly, while the books are not in their
collection, I occasionally requested them via the Interlibrary Loan
service or recommended IU libraries to purchase them.

Except for the financial support via employment and award, I have
also supported by my family, especially my father Cheng-Tien Tai (戴振沺).
He always thinks one-step ahead and cares about my future more than
I do. Discussion with other family members and friends also helped
me to compose this dissertation, to do research, or to live better
in general, and I want to thank the following people which haven't
been listed previously:\footnote{For some of them, I only provide their Chinese names for two reasons.
First, I might only know their Chinese names. Second, there are many
ways to translate a name to English, but I don't know which one they
choose.}

\begin{itemize}

>>>>

%% LyX 2.3.6.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,english]{article}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{verbatim}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{rotfloat}
\usepackage{setspace}
\usepackage[all]{xy}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newlength{\lyxlabelwidth}      % auxiliary length

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{csquotes}
\usepackage{comment}
\usepackage{framed}
\usepackage{mathrsfs}
\usepackage{setspace}
\usepackage{dsfont}
%\usepackage[AutoFallBack=true]{xeCJK}
%\setCJKmainfont[FallBack=Batang]{SimSun}
% \usepackage{xeCJK}
% \setCJKmainfont{SimSun}

@@

%\usepackage[basic]{complexity}
%\newcommand{\lang}[1]{{\ensuremath{\textsf{#1}}}}
%\newcommand{\newlang}[2]{\newcommand{#1}{\lang{#2}}}
%\newlang{\usat}{UNIQUE-SAT}
\newcommand{\ComplexityFont}[1]{%
{\ensuremath{\textsf{#1}}}%extra {} makes everyone happy.
}
\newcommand{\NP}{\ComplexityFont{NP}}

\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{prop}{Proposition}%[chapter]
\newtheorem{thm}{Theorem}%[chapter]
\newtheorem{lemma}{Lemma}%[chapter]
\newtheorem{case}{Case}
\newtheorem{corollary}{Corollary}%[chapter]

\theoremstyle{definition}
\newtheorem{definition}{Definition}%[chapter]
\newtheorem{question}{Question}%[chapter]
\newtheorem*{method}{Method}
\newtheorem{example}{Example}%[chapter]
\newtheorem{postulate}{Postulate}%[chapter]

@@

%\usepackage{unicode-math}
\RequirePackage{ifthen}
\@ifpackageloaded{unicode-math}{}{
 \newcommand{\symbf}{\mathbf}
 \newcommand{\symrm}{\mathrm}
 \newcommand{\symcal}{\mathcal}
 \newcommand{\symbb}{\mathbb}
}
\newcommand{\muB}{\ensuremath{\mu^{\symrm{B}}}}
\newcommand{\muC}{\ensuremath{\mu^{\symrm{C}}}}
\newcommand{\muF}{\ensuremath{\mu^{\symrm{F}}}}
\newcommand{\symrmL}{\ensuremath{\symrm{L}}}
\newcommand{\symrmR}{\ensuremath{\symrm{R}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\symrmL{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\symrmR{#1}}}}
\newcommand{\barmuD}{\ensuremath{\bar{\mu}^{\symrm{D}}}}
\newcommand{\events}{\ensuremath{\symcal{E}}}
\newcommand{\eventsC}{\ensuremath{\events_{\symrm{C}}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\symbb{F}_{{#1}}}
\newcommand{\Fp}{\Fpx{p}}
\newcommand{\Fppx}[1]{\Fpx{{#1}^2}}
\newcommand{\Fpp}{\Fppx{p}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\symbf{#1}}}
\newcommand{\mathReal}{\symbb{R}}
\newcommand{\mathComplex}{\symbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\symbf{S}^{#1}}
\newcommand{\R}{\mathReal}
\newcommand{\rme}{\symrm{e}}
\newcommand{\rmi}{\symrm{i}}
\newcommand{\rmd}{\symrm{d}}
\newcommand{\CP}[1]{\mathComplex\symbf{P}^{#1}} % Complex projective space
\newcommand{\DCP}[1]{\symbf{D}\mathComplex\symbf{P}^{#1}} % Discrete complex projective
\def\fh{\mathfrak{h}}
\newcommand{\uf}{U_{\!f}}
\newcommand{\scalarPlus}{+}
\newcommand{\boolt}{\textsf{Bool}} 
\newcommand{\bfalse}{\texttt{\textbf{false}}}
\newcommand{\btrue}{\texttt{\textbf{true}}}
\newcommand{\dotprod}{dot product}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\newcommand{\Psibar}{\overline{\Psi}}
\newcommand{\cardp}[2]{#1 \, \slash\!\! \slash \, #2}
\def\round{\mathop{\rm round}\nolimits}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\Hilb}{\symcal{H}}
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\def\C{{\symbb{C}}}
\newcommand{\ultramodular}{\symcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\frameL}[1][]{\ensuremath{f^{L{#1}}}}
\newcommand{\frameR}[1][]{\ensuremath{f^{R{#1}}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\nb}{\nolinebreak[3] }
\NewDocumentCommand{\jthSubsystem}{m O{} m}{{#1}_{#2}^{#3}}

@@

%\department{Department of Mathematics}
%\department{Department of Computer Science}

\usepackage{datetime2}
\usepackage{datetime2-calc}
%\monthGranted{June}%\DTMmonthname{\month}}
%\yearGranted{\the\year}

@@

\renewcommand{\tabularxcolumn}[1]{m{#1}}
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{\widthof{#1}}}

@@

\makeatother

\usepackage{babel}
\usepackage[style=numeric-comp,maxbibnames=10,doi=false,isbn=false,url=false]{biblatex}
\addbibresource{prop.bib}
\begin{document}
\title{Classical probability measures as pullback of quantum probability
measures?}

\maketitle
As discussed in Sec.~\ref{subsec:Classical-and-Quantum}, the quantum
probability spaces could be thought as a big space glued by ``local''
classical probability space defined by each orthonormal basis~$\Omega$
according to the map $\varphi:2^{\Omega}\rightarrow\events$ defined
in Eq.~(\ref{eq:pullback-function}) which sends a classical event
$E\in2^{\Omega}$ to a quantum event $\varphi\left(E\right)\in\events$.
Then, for any quantum probability measure $\mu\colon\events\rightarrow\left[0,1\right]$,
the function $\left(\lambda E.\mu\left(\varphi\left(E\right)\right)\right)\colon2^{\Omega}\rightarrow\left[0,1\right]$
defined by precomposition is a classical probability measure. Since
that ``precomposition'' is called ``pullback'' in differential
geometry~\cite{wiki:Pullback}\footnote{Whether this kind of ``pullback'' is categorical will need further
investigate.}, we will use their notation and write the precomposition $\left(\lambda E.\mu\left(\varphi\left(E\right)\right)\right)$
as $\lambda E.\left(\varphi^{*}\mu\right)\left(E\right)$, i.e., $\left(\varphi^{*}\mu\right)\left(E\right)=\mu\left(\varphi\left(E\right)\right)$,
and $\varphi^{*}\mu\colon2^{\Omega}\rightarrow\left[0,1\right]$ the
pullback of $\mu$ by $\varphi$, and summarize how ``pullback''
preserves their structures in Table~\ref{tab:Comparison}. 

@@

\noindent \centering{}\caption{\label{tab:Comparison}Comparison between the classical and quantum
measurement processes with real-valued probability for finite dimensional
cases}
%% ERT block 1 (before LyX tabular)

@@

\begin{tabular}{YYYY}

@@

Name & Classical & Quantum & How does the structure preserve?\tabularnewline

@@

Event & $E\in2^{\Omega}$ & $\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\in\events$ & Eqs.~(\ref{eq:pullback-properties-basic}) and (\ref{eq:pullback-properties})\tabularnewline

@@

Probability measure & $\begin{array}{cccc}
\varphi^{*}\mu\colon & 2^{\Omega} & \rightarrow & \left[0,1\right]\\
 & E & \mapsto & \mu\left(\varphi\left(E\right)\right)
\end{array}$ & $\begin{array}{cccc}
\mu\colon & \events & \rightarrow & \left[0,1\right]\\
 & P & \mapsto & \mu\left(P\right)
\end{array}$ & Eqs.~(\ref{eq:QuantumProbability}) and (\ref{eq:QuantumProbability-Addition})\tabularnewline

@@

@@

Possible measured values & $\lambda_{i}\in\mathbb{R}$ & $\lambda_{i}\in\mathbb{R}$ & Equal\tabularnewline

@@

The event to measure $\lambda_{i}$ & $E_{i}\in2^{\Omega}$ & $\varphi\left(E_{i}\right)\in\events$ & Mapped by $\varphi$\tabularnewline

@@

The probability that the measured value is $\lambda_{i}$ & $\left(\varphi^{*}\mu\right)\left(E_{i}\right)=\mu\left(\varphi\left(E_{i}\right)\right)\in\left[0,1\right]$ & $\mu\left(\varphi\left(E_{i}\right)\right)\in\left[0,1\right]$ & Equal\tabularnewline

@@

Summarize the measurement process as & random variable $\varphi^{*}\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}\mathbf{1}_{E_{i}}$ & observable $\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}\varphi\left(E_{i}\right)$ & \tabularnewline

@@

Expectation value & $\int\left(\varphi^{*}\mathbf{O}\right)\rmd\left(\varphi^{*}\mu\right)=\sum_{i=0}^{D-1}\lambda_{i}\left(\varphi^{*}\mu\right)\left(E_{i}\right)$ & $\expval{\mathbf{O}}_{\mu}=\sum_{i=0}^{D-1}\lambda_{i}\mu\left(\varphi\left(E_{i}\right)\right)$ & Equal, and proved in Lemma~\ref{lemma:real-expectation-value-between-classical-quantum}\tabularnewline
Probability measure could be simplified as & probability mass function $m:\Omega\rightarrow[0,1]$ such that $\left(\varphi^{*}\mu\right)(E)=\sum_{\omega\in E}m(\omega)$ & mixed state $\rho=\sum_{j=1}^{N}q_{j}\proj{\Phi_{j}}$ such that $\mu\left(P\right)=\Tr\left(\rho P\right)$ & \tabularnewline
 & Need a notation for $\varphi^{*}\mu\mapsto m$ to keep going & Need a notation for $\mu\mapsto\rho$ to keep going & We may draw commuting diagram (?) here\tabularnewline
The probability that the measured value is $\lambda_{i}$ &  &  & \tabularnewline
Expectation value &  &  & \tabularnewline
Post-measurement state &  &  & \tabularnewline

@@

We can then extend this table to compare between the classical and
quantum measurement processes to get real numbers. It is clear that
each step of the measurement process is exactly equal or just mapped
by $\varphi$ except we summarize classical measurement process as
random variable, but summarize quantum measurement process as an observable.
Notice that after we fix an orthonormal basis~$\Omega$, $\varphi$
is actually an invertible map so that given an observable $\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}P_{i}$,
we can define an unique random variable $\sum_{i=0}^{D-1}\lambda_{i}\mathbf{1}_{\varphi^{-1}\left(P_{i}\right)}$
corresponding to $\mathbf{O}$. Therefore, we call this random variable
as the pullback of $\mathbf{O}$ by $\varphi$, and denoted by $\varphi^{*}\mathbf{O}$.\footnote{We might also choose the other way around so that each random variable
has a push forward observable. This is better in the sense that we
don't have the ambiguity that $\varphi^{*}\mathbf{O}$ is might be
different for different orthonormal basis, but every theorem for IVPMs
starts from the quantum part with an observable, so we need the ambiguous
$\varphi^{*}\mathbf{O}$ for each different orthonormal basis anyway.} According to Gleason's theorem, the table could be extended to simplify
the formula computing expectation values and the the post-measurement-state
postulates after we define a notation to send probability measures
to states and the other way around.

By following the same process, Sec.~\ref{sec:Interval-Uncertainty}
showed the corresponding results for interval-valued probability.
For any QIVPM $\bar{\mu}\colon\events\rightarrow\mathscr{I}$, the
function $\varphi^{*}\bar{\mu}\colon2^{\Omega}\rightarrow\mathscr{I}$
defined by $\left(\varphi^{*}\bar{\mu}\right)\left(E\right)=\bar{\mu}\left(\varphi\left(E\right)\right)$
is a classical IVPM.
\begin{sidewaystable}
\noindent \centering{}\caption{\label{tab:Comparison-interval}Comparison between the classical and
quantum measurement processes with interval-valued probability for
finite dimensional cases}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{YYYY}
\toprule 
Name & Classical & Quantum & How does the structure preserve?\tabularnewline
\midrule
\midrule 
Probability measure & $\begin{array}{cccc}
\varphi^{*}\bar{\mu}\colon & 2^{\Omega} & \rightarrow & \mathscr{I}\\
 & E & \mapsto & \bar{\mu}\left(\varphi\left(E\right)\right)
\end{array}$ & $\begin{array}{cccc}
\bar{\mu}\colon & \events & \rightarrow & \mathscr{I}\\
 & E & \mapsto & \bar{\mu}\left(E\right)
\end{array}$ & \tabularnewline
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{sidewaystable}
